Effect of Beam Chromaticity on Foregrounds in Wide-Field Measurements of Redshifted 21~cm Power Spectra

### 2016-03-01 ForegroundSimulations/ms.tex
Rearrange section and subsection titles
From 508edb8 (restructure):
==========================

In eor.bib,

(*) Add an entry for Harris (1978) on window functions

In ms.tex,

(*) Rearrange section and subsection titles
(*) Add couple of paragraphs on new spectral window function

\subsection{Directional Chromaticity of Power Pattern}\label{sec:directional-chromaticity}

Fig.~\ref{fig:off-axis-beam-chromaticity} demonstrates the beyond-the-horizon spillover even from a flat-spectrum point source at different off-axis angles. The three panels correspond to delay spectra from different positions of the point source -- 0\arcdeg (left), 45\arcdeg (middle) and 89\arcdeg (right) -- from the zenith. The phase centers are located at the source positions and hence the delay spectra are centered on $\tau = 0$~ns. The response of the simulated dish power pattern (dashed line) is comapred with that from an {\it Airy} pattern resulting from a nominal uniformly illuminated circular disk (solid line). The gray vertical lines correspond to the horizon limits at $\pm 48.7$~ns for a 14.6~m antenna spacing. In all these cases, a {\it Blackman-Harris} window function has been applied along the spectral axis.

\begin{figure*}[htb]
\centering
\includegraphics[width=\linewidth]{off-axis_point_source_beam_chromaticity.eps}
\caption{Chromaticity of antenna power pattern at directions off-axis. }
\label{fig:off-axis-beam-chromaticity}
\end{figure*}

At 0\arcdeg off-axis, the {\it Airy} pattern has no spectral variation except that introduced by the {\it Blackman-Harris} window function and thus reproduces its Fourier response centered at $\tau = 0$~ns whereas the simulated power pattern is found to exhibit some spectral variation giving rise to the wings on either side of $\tau = 0$~ns. As off-axis angle increases to 45\arcdeg, both the power patterns clearly exhibit chromaticity. The simulated pattern not only has a higher magnitude inside the horizon limits (by a factor $\sim 10$) but also exhibits higher chromaticity (by many orders of magnitude) in modes beyond the horizon relative to that of an {\it Airy} pattern. In contrast, at 89\arcdeg off-axis, not only does the simulated pattern have higher chromaticity in high delay modes but is also two orders of magnitude higher in amplitude inside the horizon limits relative to the nominal {\it Airy} pattern.

It is important to note that in a generic scenario where visibilities are not phased to a specific foreground location, the chromaticity of the power pattern will imprint itself on location of the foreground objects in delay space and will give rise to significant spillover especially from foregrounds near the horizon.

We generalize the analysis above by computing the delay power spectrum of the power pattern along its spectral axis in every direction in an effort to study its performance as a function of angle on the sky. 
\begin{align}
  A^\prime(l,m,\tau) &= \int A(l,m,f)\,e^{-j 2\pi f\tau} \,\dif f
\end{align}

We define the directional chromaticity of antenna power pattern as:
\begin{align}
  C(l,m) &= \frac{ \int\limits_{|\tau|>60\,\textrm{ns}} |A^\prime|^2 \,\dif\tau }{\int\limits_{|\tau|>60\,\textrm{ns}} \dif\tau}
\end{align}
where, the limit $|\tau|>60$~ns was chosen to represent a region well beyond the horizon limits ($\pm 48.7$~ns) for a 14.6~m antenna spacing. It thus measures the average power arising out of chromaticity in each delay mode beyond $|\tau|>60$~ns.

Fig.~\ref{fig:directional-beam-chromaticity} shows the directional chromaticity, $C(l,m)$, of the power patterns of an {\it Airy} disk for reference (left) and a simulated disk (middle) on an orthographic projection onto direction cosines. Their color scale is shown at the top for reference. The chromaticity of the {\it Airy} pattern is symmetric and always below 500. On the other hand,  a simulated disk exhibits higher levels of chromaticity in the range 100--50,000. The ratio of their chromaticities is shown on the right with corresponding color scale on its right. The simulated disk is more chromatic than the nominal {\it Airy} disk by a factor between 100 and up to 10000. However, chromaticity in regions near the horizon which map to the horizon limits in the delay spectrum and have the most significant impact on spillover into the {\it EoR window} in the simulated disk are higher than that of an {\it Airy} disk only by factors $\lesssim 4$. Thus the simulated disk for HERA achieves reasonable control on the chromaticity near the horizon despite the very high levels near the center.

Also shown are the tracks (black dots) of point sources brighter than 50~Jy in the LST range 0--12~hrs. The solid circle near the bottom denotes the south celestial pole. It is noted that there are a few bright point sources that spend a majority of time in this LST range near the northern and southern horizons which will contaminate the measured visibilities on north-south antenna spacings. Simulations of the antenna power pattern must take the bright foreground locations, especially near the horizon, into account.

\begin{figure*}[htb]
\centering
\includegraphics[width=\linewidth]{directional_high_delay_average_in_beam.eps}
\caption{All-sky directional chromaticity of antenna power pattern. }
\label{fig:directional-beam-chromaticity}
\end{figure*}

@@

>>>>

\subsection{Effects on Foreground Contamination}\label{sec:effects-fgdps}

In the estimation of power spectrum, specific choices for spectral weighting function, $W(f)$, has been found to be effective in reducing foreground contamination by many orders of magnitude \citep{thy13} and is regularly used in EoR data analysis \citep{par12a}. For instance, a {\it Blackman-Harris} function \citep{har78} has a dynamic range of $\sim$100 -- 120~dB in delay power spectrum and a reduced effective bandwidth, where, we define effective bandwidth as:
\begin{align}
  B_\textrm{eff} &= \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f.
\end{align}
For a {\it Blackman-Harris} window function denoted by $W_\textrm{BH}(f)$, $B_\textrm{eff} \approx B/2$. Hence, the window spans a total of $2N$ frequency channels each of width $\Delta f$ to have $B_\textrm{eff} \approx N\,\Delta f$.

Considering that the dynamic range required to suppress foreground contamination in the {\it EoR window} may be even higher than that provided by a {\it Blackman-Harris} window function, we use a modified version given by the convolution of a {\it Blackman-Harris} window with itself:
\begin{align}
  W(f) &= W_\textrm{BH}(f) \ast W_\textrm{BH}(f).
\end{align}
In delay spectrum, this window has a response that is obtained by squaring of the {\it Blackman-Harris} window response and thus improves the dynamic range further in the response by another 100 -- 120~dB.

\subsection{Constraints on Reflections in the Instrument}\label{sec:constraints-reflectometry}

% \subsection{Directional Chromaticity of Power Pattern}\label{sec:directional-chromaticity}

% Fig.~\ref{fig:off-axis-beam-chromaticity} demonstrates the beyond-the-horizon spillover even from a flat-spectrum point source at different off-axis angles. The three panels correspond to delay spectra from different positions of the point source -- 0\arcdeg (left), 45\arcdeg (middle) and 89\arcdeg (right) -- from the zenith. The phase centers are located at the source positions and hence the delay spectra are centered on $\tau = 0$~ns. The response of the simulated dish power pattern (dashed line) is comapred with that from an {\it Airy} pattern resulting from a nominal uniformly illuminated circular disk (solid line). The gray vertical lines correspond to the horizon limits at $\pm 48.7$~ns for a 14.6~m antenna spacing. In all these cases, a {\it Blackman-Harris} window function has been applied along the spectral axis.

% \begin{figure*}[htb]
% \centering
% \includegraphics[width=\linewidth]{off-axis_point_source_beam_chromaticity.eps}
% \caption{Chromaticity of antenna power pattern at directions off-axis. }
% \label{fig:off-axis-beam-chromaticity}
% \end{figure*}

% At 0\arcdeg off-axis, the {\it Airy} pattern has no spectral variation except that introduced by the {\it Blackman-Harris} window function and thus reproduces its Fourier response centered at $\tau = 0$~ns whereas the simulated power pattern is found to exhibit some spectral variation giving rise to the wings on either side of $\tau = 0$~ns. As off-axis angle increases to 45\arcdeg, both the power patterns clearly exhibit chromaticity. The simulated pattern not only has a higher magnitude inside the horizon limits (by a factor $\sim 10$) but also exhibits higher chromaticity (by many orders of magnitude) in modes beyond the horizon relative to that of an {\it Airy} pattern. In contrast, at 89\arcdeg off-axis, not only does the simulated pattern have higher chromaticity in high delay modes but is also two orders of magnitude higher in amplitude inside the horizon limits relative to the nominal {\it Airy} pattern.

% It is important to note that in a generic scenario where visibilities are not phased to a specific foreground location, the chromaticity of the power pattern will imprint itself on location of the foreground objects in delay space and will give rise to significant spillover especially from foregrounds near the horizon.

% We generalize the analysis above by computing the delay power spectrum of the power pattern along its spectral axis in every direction in an effort to study its performance as a function of angle on the sky. 
% \begin{align}
%   A^\prime(l,m,\tau) &= \int A(l,m,f)\,e^{-j 2\pi f\tau} \,\dif f
% \end{align}

% We define the directional chromaticity of antenna power pattern as:
% \begin{align}
%   C(l,m) &= \frac{ \int\limits_{|\tau|>60\,\textrm{ns}} |A^\prime|^2 \,\dif\tau }{\int\limits_{|\tau|>60\,\textrm{ns}} \dif\tau}
% \end{align}
% where, the limit $|\tau|>60$~ns was chosen to represent a region well beyond the horizon limits ($\pm 48.7$~ns) for a 14.6~m antenna spacing. It thus measures the average power arising out of chromaticity in each delay mode beyond $|\tau|>60$~ns.

% Fig.~\ref{fig:directional-beam-chromaticity} shows the directional chromaticity, $C(l,m)$, of the power patterns of an {\it Airy} disk for reference (left) and a simulated disk (middle) on an orthographic projection onto direction cosines. Their color scale is shown at the top for reference. The chromaticity of the {\it Airy} pattern is symmetric and always below 500. On the other hand,  a simulated disk exhibits higher levels of chromaticity in the range 100--50,000. The ratio of their chromaticities is shown on the right with corresponding color scale on its right. The simulated disk is more chromatic than the nominal {\it Airy} disk by a factor between 100 and up to 10000. However, chromaticity in regions near the horizon which map to the horizon limits in the delay spectrum and have the most significant impact on spillover into the {\it EoR window} in the simulated disk are higher than that of an {\it Airy} disk only by factors $\lesssim 4$. Thus the simulated disk for HERA achieves reasonable control on the chromaticity near the horizon despite the very high levels near the center.

% Also shown are the tracks (black dots) of point sources brighter than 50~Jy in the LST range 0--12~hrs. The solid circle near the bottom denotes the south celestial pole. It is noted that there are a few bright point sources that spend a majority of time in this LST range near the northern and southern horizons which will contaminate the measured visibilities on north-south antenna spacings. Simulations of the antenna power pattern must take the bright foreground locations, especially near the horizon, into account.

% \begin{figure*}[htb]
% \centering
% \includegraphics[width=\linewidth]{directional_high_delay_average_in_beam.eps}
% \caption{All-sky directional chromaticity of antenna power pattern. }
% \label{fig:directional-beam-chromaticity}
% \end{figure*}

@@

\section{EoR Sensitivity}\label{sec:eor-sensitivity}
